Local wind regime induced by giant linear dunes: comparison of ERA5-Land reanalysis with surface measurements

### 2022-07-04 Paper/main.tex
A few modifications on Philippe's round

CITE \citet{Chand22}

@@

%---------------------------------------------------------------------------------------
\section{Discussion and conclusion}
The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms. As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}. These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'~\citep{Chand22}. Whilst differences between large and small scale dune patterns are observed ubiquitously, they are usually attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}. Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes governed by the same dune growth mechanism (elongating mode). Figure~\ref{Fig9} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig9}). These predictions require to specify the threshold of aeolian sand transport. Importantly, its value (a shear velocity estimated $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) is reached in the winds which are deflected -- recall that the larger winds are not deflected, see Fig.~\ref{Fig6}. The feedback of the giant dunes on the wind described in this study thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date. These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to that currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}. Further work is needed to investigate these processes in more detail, including measurements of sediment transport and flow on the top of dunes.

@@

%---------------------------------------------------------------------------------------
\begin{acknowledgements}
We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.

@@

We thank B. Andreotti for useful discussions. Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.

@@

%---------------------------------------------
\begin{figure}[p]
\centering
\includegraphics[scale=1]{Figures/Figure3_supp.pdf}
\caption{Calibration of hydrodynamic roughness. The parameter $\delta$ \eqref{metric_roughness} quantifying the difference between local and predicted winds is shown in color scale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations. The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively. The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
\label{Fig3_supp}
\end{figure}
%---------------------------------------------

>>>>

@@

%---------------------------------------------------------------------------------------
\section{Discussion and conclusion}
The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms. As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}. These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'~\citep{Chand22}. Whilst differences between large and small scale dune patterns are observed ubiquitously, they are usually attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}. Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes governed by the same dune growth mechanism (elongating mode). Figure~\ref{Fig9} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig9}). These predictions require to specify the threshold of aeolian sand transport. Importantly, its value (a shear velocity estimated at $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) can be reached in periods during which winds are deflected (recall that the stronger winds, responsible for most of the sediment transport and associated dune morphodynamics, are not deflected -- see Fig.~\ref{Fig6}). The feedback of the giant dunes on the wind described in this study, through wind delection and attenuation, thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date. These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to that currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}. Further work is needed to investigate these processes in more detail, including measurements of sediment transport and flow on the top of dunes.

@@

%---------------------------------------------------------------------------------------
\begin{acknowledgements}
We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python. We also thank B. Andreotti and A. Gunn for useful discussions.

@@

Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.

@@

%---------------------------------------------
\begin{figure}[p]
\centering
\includegraphics[scale=1]{Figures/Figure3_supp.pdf}
\caption{Calibration of hydrodynamic roughness. The parameter $\delta$ \eqref{metric_roughness} quantifying the difference between local and predicted winds is shown in color scale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations. The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively. The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5-Land}} = 1$~mm.}
\label{Fig3_supp}
\end{figure}
%---------------------------------------------
